\textbf{{\Large {System of dual functions in the reduction of affine Lie algebra representations}}}

### 2010-05-21 AffineLieAlgebras/DualFunctions/article.tex
Some files added, pdf regenerated

\begin{document}
\thispagestyle{empty}

\begin{center}
{\Large \bf SAINT-PETERSBURG STATE\\[0.5cm] UNIVERSITY}
\end{center}
\vskip 1 cm
\begin{flushright}
{\large \bf SPbU-IP-09-08}
\end{flushright}
\hfill {}
\vskip 1 cm
\begin{center}
{\Huge \bf Recursive algorithms, branching coefficients \\ for affine algebras and applications\\[0.5cm]
}
\end{center}
\vskip 1.0 cm
\begin{center}
{\Large \bf
Vladimir Lyakhovsky\\[2mm]
Anton Nazarov}
\end{center}
\vfill
\begin{center}
{\large Department of Theoretical Physics\\
}
\end{center}
\vfill
\begin{center}
SAINT-PETERSBURG \\
2009
\end{center}

\newpage

\pagenumbering{arabic}
\title{\textbf{{\Large {Recursive algorithms, branching coefficients affine algebras and applications}}}}
\author{Vladimir Lyakhovsky \thanks{ Supported by
 RFFI grant N 09-01-00504 and the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:lyakh1507@nm.ru \\
[5mm] Anton Nazarov \thanks{ Supported by
the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:antonnaz@gmail.com
}
\maketitle

\begin{abstract}
  Recurrent relations for branching coefficients in affine Lie algebras highest weight modules are studied. The decomposition algorithm for  integrable highest weight modules reduction based on the injection fan technique is adopted to the situation where the Weyl denominator becomes singular with respect to a reductive  subalgebra. We study some modifications of the injection fan technique and demonstrate that it is possible to define the "subtracted fans" that play the role similar to the original one. The possible applications of subtracted fans in CFT models are considered.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

The problem of reduction of a Lie algebra representation to a  subalgebra is studied for several decades and has various applications in physics. In the context of finite-dimensional algebras it is important for the study of great unification models whilst the branching problem for affine Lie algebras emerges in conformal field theory, for example, in the construction of the modular-invariant partition functions \cite{difrancesco1997cft}.

There exist several approaches to the computation of the branching coefficients. Some of them use the BGG
resolution \cite{bernstein1975differential} (for Kac-Moody algebras the algorithm is described in
\cite{kac1990idl},\cite{wakimoto2001idl}), the Schure function series \cite{fauser2006new}, the BRST
cohomology \cite{hwang1994general}, Kac-Peterson formulas \cite{kac1990idl,quella2002branching} or the
combinatorial methods applied in \cite{feigin707principal}.

Usually only the embedding of the maximal reductive subalgebras is considered since the case of non-maximal subalgebra can be obtained using the chain of maximal injections. In this paper we study the recurrent properties for branching coefficients which generalise the recurrent relations obtained earlier (see the paper \cite{ilyin812pbc}and the references therein) to the cases of non-maximal reductive subalgebras. The result is formulated by an introduction of a new injection fan called "subtracted fan". Using this new tools we formulate simple and explicit algorithm for the computation of branching coefficients which is applicable to the non-maximal subalgebras of finite-dimensional and affine Lie algebras.

We show that our algorithm can be used in a study of conformal embeddings where the central charge of the conformal field theory is preserved, and the computations are simplified by taking into account some  physical limitations.

The paper is organised as follows. In the next subsection  we fix the notations used throughout the paper. In the Section \ref{sec:recurr-form-branch} we derive the subtracted recurrent formula for anomalous branching coefficients and describe the decomposition algorithm for integrable highest weight modules of algebra $\mathfrak{g}$ with respect to a reductive subalgebra $\mathfrak{a}$ (subsection \ref{sec:algorithm}). In the next Section \ref{sec:examples} we present several examples and discuss some applications in CFT models (s. \ref{sec:phys-appl}). We conclude the paper with a review of obtained results. Possible future developments are discussed (s. \ref{sec:conclusion}).

\subsection{Notation}
\label{sec:notation}

Consider affine Lie algebras $\frak{g}$ and $\frak{a}$ with the
underlying finite-dimensional subalgebras $\go$ and $%
\ao$ and an injection $\frak{a}\longrightarrow \frak{g%
}$ such that $\frak{a}$ is a reductive subalgebra $\frak{a\subset g}$ with
correlated root spaces: $\frak{h}_{\frak{a}}^{\ast }\subset \frak{h}_{\frak{g%
}}^{\ast }$ and $\frak{h}_{\ao}^{\ast }\subset \frak{h%
}_{\go}^{\ast }$\ .

We use the following notations adopted from the paper \cite{ilyin812pbc}.

$L^{\mu }$\ $\left( L_{\frak{a}}^{\nu }\right) $\ --- the integrable module
of $\frak{g}$ with the highest weight $\mu $\ ; (resp. integrable $\frak{a}$
-module with the highest weight $\nu $ );

$r$ , $\left( r_{\frak{a}}\right) $ --- the rank of the algebra $\frak{g}$ $%
\left( \text{resp. }\frak{a}\right) $ ;

$\Delta $ $\left( \Delta _{\frak{a}}\right) $--- the root system; $\Delta
^{+} $ $\left( \text{resp. }\Delta _{\frak{a}}^{+}\right) $--- the positive
root system (of $\frak{g}$ and $\frak{a}$ respectively);

$\mathrm{mult}\left( \alpha \right) $ $\left( \mathrm{mult}_{\frak{a}}\left(
\alpha \right) \right) $ --- the multiplicity of the root $\alpha$ in $\Delta
$ (resp. in $\left( \Delta _{\frak{a}}\right) $);

$\co{\Delta}$ , $\left( \co{\Delta _{\frak{a}}}%
\right)$ --- the finite root system of the subalgebra $\co{%
\frak{g}}$ (resp. $\co{\frak{a}}$);
$\Theta$, $(\Theta_{\mathfrak{a}})$ --- the highest root of the algebra $\mathfrak{g}$ (resp. subalgebra $\mathfrak{a}$);

$\mathcal{N}^{\mu }$ , $\left( \mathcal{N}_{\frak{a}}^{\nu }\right) $ --- the
weight diagram of $L^{\mu }$ $\left( \text{resp. }L_{\frak{a}}^{\nu }\right)
$ ;

$W$ , $\left( W_{\frak{a}}\right) $--- the corresponding Weyl group;

$C$ , $\left( C_{\frak{a}}\right) $--- the fundamental Weyl chamber;

$\bar{C}, \left(\bar{C_{\mathfrak{a}}}\right)$ --- the closure of the fundamental Weyl chamber;

$\rho $\ , $\left( \rho _{\frak{a}}\right) $\ --- the Weyl vector;

$\epsilon \left( w\right) :=\det \left( w\right) $ ;

$\alpha _{i}$ , $\left( \alpha _{\left( \frak{a}\right) j}\right) $ --- the $i
$-th (resp. $j$-th) basic root for $\frak{g}$ $\left( \text{resp. }\frak{a}%
\right) $; $i=0,\ldots ,r$ ,\ \ $\left( j=0,\ldots ,r_{\frak{a}}\right) $;

$\delta $ --- the imaginary root of $\frak{g}$ (and of $\frak{a}$ if any);

$\alpha _{i}^{\vee }$ , $\left( \alpha _{\left( \frak{a}\right) j}^{\vee
}\right) $--- the basic coroot for $\frak{g}$ $\left( \text{resp. }\frak{a}%
\right) $ , $i=0,\ldots ,r$ ;\ \ $\left( j=0,\ldots ,r_{\frak{a}}\right) $;

$\co{\xi }$ , $\co{\xi _{\left( \frak{a}\right) }}$
--- the finite (classical) part of the weight $\xi \in P$ , $\left( \text{%
resp. }\xi _{\left( \frak{a}\right) }\in P_{\frak{a}}\right) $;

$\lambda =\left( \co{\lambda };k;n\right) $ --- the
decomposition of an affine weight indicating the finite part $\co{\lambda }$, level $k$ and grade $n$\ .

$P$ $\left( \text{resp. } P_{\frak{a}}\right) $ \ --- the weight lattice;

$M \left( \text{resp. }M_{\frak{a}}\right) :=$

\noindent $=\left\{
\begin{array}{c}
\sum_{i=1}^{r}\mathbf{Z}\alpha _{i}^{\vee }\text{ }\left( \text{resp. }%
\sum_{i=1}^{r}\mathbf{Z}\alpha _{\left( \frak{a}\right) i}^{\vee }\right)
\text{for untwisted algebras or }A_{2r}^{\left( 2\right) }, \\
\sum_{i=1}^{r}\mathbf{Z}\alpha _{i}\text{ }\left( \text{resp. }\sum_{i=1}^{r}%
\mathbf{Z}\alpha _{\left( \frak{a}\right) i}\right) \text{for }A_{r}^{\left(
u\geq 2\right) }\text{ and }A\neq A_{2r}^{\left( 2\right) },
\end{array}
\right\} ;$
$\Psi ^{\left( \mu \right) }:=\sum\limits_{w\in W}\epsilon (w)e^{w\circ (\mu +\rho )-\rho }$ --- the singular weight element for the $\frak{g}$-module $L^{\mu }$;
$\Psi _{\left( \frak{a}\right) }^{\left( \nu \right) }:=\sum\limits_{w\in W_{\frak{a}}}\epsilon (w)e^{w\circ (\nu +\rho
_{_{\frak{a}}})-\rho _{_{\frak{a}}}}$ --- the corresponding singular weight
element for the $\frak{a}$-module $L_{\frak{a}}^{\nu }$;

$\widehat{\Psi ^{\left( \mu \right) }}$ $\left( \widehat{\Psi _{\left( \frak{%
a}\right) }^{\left( \nu \right) }}\right) $ --- the set of singular weights $%
\xi \in P$ $\left( \text{resp. }\in P_{\frak{a}}\right) $ for the module $%
L^{\mu }$ $\left( \text{resp. }L_{\frak{a}}^{\nu }\right) $ with the
coordinates $\left( \co{\xi },k,n,\epsilon \left( w\left( \xi
\right) \right) \right) \mid _{\xi =w\left( \xi \right) \circ (\mu +\rho
)-\rho },$ (resp. $\left( \co{\xi },k,n,\epsilon \left(
w_{a}\left( \xi \right) \right) \right) \mid _{\xi =w_{a}\left( \xi \right)
\circ (\nu +\rho _{a})-\rho _{a}}$ ), (this set is similar to $P_{\mathrm{%
nice}}^{\prime }\left( \mu \right) $ in \cite{wakimoto2001idl})

$m_{\xi }^{\left( \mu \right) }$ , $\left( m_{\xi }^{\left( \nu \right)
}\right) $ --- the multiplicity of the weight $\xi \in P$ \ $\left( \text{%
resp. }\in P_{\frak{a}}\right) $ in the module $L^{\mu }$ , (resp. $\xi \in
L_{\frak{a}}^{\nu } $);

$ch\left( L^{\mu }\right) $ $\left( \text{resp. }ch\left( L_{\frak{a}}^{\nu
}\right) \right) $--- the formal character of $L^{\mu }$ $\left( \text{resp. }%
L_{\frak{a}}^{\nu }\right) $;

$ch\left( L^{\mu }\right) =\frac{\sum_{w\in W}\epsilon (w)e^{w\circ (\mu
+\rho )-\rho }}{\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{%
\mathrm{{mult}\left( \alpha \right) }}}=\frac{\Psi ^{\left( \mu \right) }}{%
\Psi ^{\left( 0\right) }}$ --- the Weyl-Kac formula.

$R:=\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{{%
mult}\left( \alpha \right) }}=\Psi ^{\left( 0\right) }\quad $

\noindent $\left( \text{resp. }R_{\frak{a}}:=\prod_{\alpha \in \Delta _{%
\frak{a}}^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{mult}_{\frak{a}}\mathrm{%
\left( \alpha \right) }}=\Psi _{\frak{a}}^{\left( 0\right) }\right) $--- the
denominator.

$  L_{\frak{g}\downarrow \frak{a}}^{\mu }=\bigoplus\limits_{\nu \in P_{\frak{a}}^{+}}b_{\nu }^{\left( \mu \right) }L_{\frak{a}}^{\nu }$ --- the reduction of the module;

$b^{(\mu)}_{\nu}$ --- the branching coefficients;

\begin{equation}
  \label{eq:21}
  \sum_{\nu \in \bar{C_{\mathfrak{a}}}}b_{\nu }^{\left( \mu \right) }\Psi _{\left( \frak{%
        a}\right) }^{\left( \nu \right) }=\sum_{\lambda \in P_{\frak{a}}}k_{\lambda
  }^{\left( \mu \right) }e^{\lambda }
\end{equation}

 $k_{\lambda}$ --- the anomalous branching coefficients;

It is important to mention that
\begin{equation}
  \label{eq:20}
  b^{(\mu)}_{\nu}=k^{(\mu)}_{\nu} \; \mbox{for} \; \nu\in \bar{C}_{\mathfrak{a}}
\end{equation}

$x_e=\frac{\left|\pi_{\mathfrak{a}} \Theta\right|^2}{\left|\Theta_{\mathfrak{a}}\right|^2}$ --- the embedding index.

\section{Recurrent relation for branching coefficients. Singularities and subtractions}
\label{sec:recurr-form-branch}

Our aim is to demonstrate that despite possible singularities arriving in the Weyl denominator
when it is projected to the subalgebra root space injection fan technique can be properly modified.
The result of such modification is that the generalized recurrent relations for anomalous branching
coefficients \eqref{eq:21} must be reformulated in the following form:
\begin{equation}
  \label{eq:1}
  \begin{array}{c}
      k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
  \sum_{\omega\in W_{\bot}\backslash W} \epsilon(\omega)\; {\rm dim}\left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right) \delta_{\xi-\gamma_0,\pi_{\mathfrak{a}}(\omega(\mu+\rho)-\rho)}+ \right.\\
\left.
+\sum_{\gamma \in
\Gamma _{\frak{a}\subset \frak{g}}}s\left( \gamma +\gamma _{0}\right) k_{\xi
+\gamma }^{\left( \mu \right) }\right)
  \end{array}
   \end{equation}
Here $\mathfrak{a}_{\bot}$ is the subalgebra described by the roots of $\mathfrak{g}$
orthogonal to the root subsystem of $\mathfrak{a}$, $W_{\bot}$ is the corresponding Weyl group,
$\Gamma_{\mathfrak{a}\subset \mathfrak{g}}$ is the set of weights in the expansion
of the denominator $\prod_{\alpha\in \Delta^{+}\setminus \Delta^{+}_{\mathfrak{a}_{\bot}}}
(1-e^{-\alpha})^{\mathrm{mult}(\alpha)-\mathrm{mult}_{\mathfrak{a}}(\alpha)}$ and $s(\gamma)$
is the coefficient of $e^{\gamma}$ in this expansion.
In the next subsection we study the situation in details and prove the validity of this relation.

In the section \ref{sec:algorithm} we  shall describe the  computational algorithm for  branching coefficients
based on this formula and present some examples.

\subsection{Proof of the recurrent relation}
\label{sec:proof}

Consider the branching of a module $L_{\frak{g}}^{\mu }$ in terms of formal characters and
projection operators $\pi_{\mathfrak{a}}$ that bring the roots of $\mathfrak{g}$ to the
weight subspace of $\mathfrak{a}$:
\begin{equation}
  \label{eq:3}
  L_{\frak{g}\downarrow \frak{a}}^{\mu }=\bigoplus\limits_{\nu \in P_{\frak{a}%
    }^{+}}b_{\nu }^{\left( \mu \right) }L_{\frak{a}}^{\nu }\quad
  \Longrightarrow\quad
  \pi_{\mathfrak{a}}(ch L^{\mu}_{\mathfrak{g}})=
  \sum_{\nu\in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu} ch L^{\nu}_{\mathfrak{a}}
\end{equation}
The Weyl-Kac character formula leads to
the equality
\begin{equation}
  \label{eq:4}
  \pi_{\mathfrak{a}}\left(\frac{\sum_{\omega\in W} \epsilon(\omega) e^{\omega(\mu+\rho)-\rho}}
  {\prod_{\alpha\in\Delta^{+}}(1-e^{-\alpha})^{\mathrm{mult}(\alpha)}}\right) =
  \sum_{\nu\in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu}
  \frac{\sum_{\omega\in W_{\mathfrak{a}}}\epsilon(\omega)
  e^{\omega(\nu+\rho_{\mathfrak{a}})-\rho_{\mathfrak{a}}}}
  {\prod_{\beta\in \Delta_{\mathfrak{a}}^{+}}(1-e^{-\beta})^{\mathrm{mult}_{\mathfrak{a}}(\beta)}}
\end{equation}

It is important to keep in mind that the projection of some of positive roots of
the algebra $\mathfrak{g}$ can be equal to zero. These roots are orthogonal to the
root space of the subalgebra $\mathfrak{a}$ embedded into the root space of the
algebra $\mathfrak{g}$. Let's denote the subset of such roots by $\Delta^{+}_{\bot}
=\left\{\alpha\in\Delta_{\mathfrak{g}}^{+}:
\forall \beta\in \Delta_{\mathfrak{a}}^{+},\; \alpha\bot\beta \right\}$.

Notice that if the set $\Delta^{+}_{\bot}$ is non-empty the Weyl reflections
corresponding to the positive roots of $\Delta^{+}_{\bot}$ generate a subgroup
$W_{\bot}$ of the Weyl group $W$. Consider any two positive roots $\alpha,\;
\beta\in \Delta^{+}_{\bot}$ and the corresponding Weyl reflections
$\omega_{\alpha},\; \omega_{\beta}\in W_{\bot}$.  Since roots of the subalgebra
$\mathfrak{a}$ are invariant under $\omega_{\alpha}, \; \omega_{\beta}$ they
are also invariant under the action of $\omega_{\gamma}=\omega_{\alpha}\cdot \omega_{\beta}$.
So the subgroup $W_{\bot}$ preserves the root system of the subalgebra $\mathfrak{a}$.

Thus we have obtained the root system $\Delta_{\bot}$ which is orthogonal to the root
system $\Delta_{\mathfrak{a}}$ and invariant with respect to  $W_{\bot}$ . This root system
can be considered as the root system of a subalgebra $\mathfrak{a}_{\bot}\subset \mathfrak{g}$.

Now we are to find out when the subset $\Delta^{+}_{\bot}$ is non-empty
and the subgroup $W_{\bot}$ and subalgebra $\mathfrak{a}_{\bot}$ are non-trivial.

If $\mathfrak{a}$ is a maximal regular subalgebra of $\mathfrak{g}$ then
the rank of $\mathfrak{a}$ is equal to the rank of $\mathfrak{g}$ and
it is clear that $\Delta^{+}_{\bot}$ is empty.
On the other hand
non-maximal regular embedding of $\mathfrak{a}$ into $\mathfrak{g}$ can be obtained
through the chain of maximal embeddings
$\mathfrak{a}\subset \mathfrak{p}_1\subset \mathfrak{p}_2\subset\dots \subset \mathfrak{g}$.
The maximal regular embeddings are constructed by the exclusion of one or two roots
from the extended Dynkin diagram of the algebra. Since this process can give us
non-connected Dynkin diagrams we can see which roots are orthogonal to the root
space of non-maximal regular subalgebra $\mathfrak{a}$.

Consider for instance regular the embedding of $A_1\subset B_2$.

The extended Dynkin diagram of $B_2$ is presented in the Figure \ref{fig:B2Dynkin}.
\begin{figure}[ph]
  \centering
  \includegraphics[width=50mm]{B2_A1_diagram.pdf}
  \caption{Extended Dynkin diagram of $B_2$ and embedding of $A_1$}
  \label{fig:B2Dynkin}
\end{figure}
Drop the central node to describe the embedding $A_1\oplus A_1\subset B_2$.
In this case we have: $\mathfrak{a}=A_1$ and $\mathfrak{a}_{\bot}=A_1$.

The simple criterion of $\Delta^{+}_{\bot}$'s non-emptiness for a regular
embedding $\mathfrak{a}\subset \mathfrak{g}$
when both $\mathfrak{a}$ and $\mathfrak{g}$ are simple can be formulated
as follows:
if the Dynkin diagram of $\mathfrak{g}$ can be split into the disconnected
diagrams of $\mathfrak{a}$ and of some subalgebras $\{\mathfrak{\bar{a}}_j\}$
then the subset $\Delta_{\bot}$ is non-empty,
subalgebra $\mathfrak{a}_{\bot}$ is non-trivial and all the
$\mathfrak{\bar{a}}_j$ are the subalgebras of $\mathfrak{a}_{\bot}$.

Note that when we study the regular embedding obtained by dropping the nodes
of the extended Dynkin diagram of the algebra $\mathfrak{g}$ and the
subalgebra $\mathfrak{a}$ is one of the connected components,
the subalgebra $\mathfrak{a}_{\bot}$ may be larger than the algebra
generated by the remaining connected components.
Consider for example the embedding of $B_2\subset B_4$ (the figure \ref{fig:dynkin}).
In this case by eliminating
the simple root $\alpha_2=e_2-e_3$ one splits the extended Dynkin diagram of $B_4$
into the diagrams of the subalgebra $\mathfrak{a}=B_2$ and that of the direct
sum $A_1 \oplus A_1 $. But the subalgebra $\mathfrak{a}_{\bot}$ is equal not to
$A_1\oplus A_1$ but to $B_2$ (the root system of $B_4$ contains not only $\alpha_2=e_2-e_3$ but also $e_2$).

Such effects are due to the fact that the subalgebras $\mathfrak{a}$ and $\mathfrak{a}_{\bot}$ must not form a direct sum in $\mathfrak{g}$.  
Consider the case of such a regular embedding $\mathfrak{a}\subset \mathfrak{g}$ where both algebras are simple and the diagram of the subalgebra $\mathfrak{a}_{\bot}$ is not a subdiagram of the extended Dynkin diagram $\mathfrak{g}$.
 Drop the subdiagram of $\mathfrak{a}$ and the node $\alpha'$ that connects it with all the remaining nodes of the diagram of $\mathfrak{g}$. Consider the remaining diagram. This diagram is the diagram of the algebra $\mathfrak{\bar{a}}$ of $\mathrm{rank}(\mathfrak{\bar{a}}) = \mathrm{rank}(\mathfrak{g})-\mathrm{rank}(\mathfrak{a})$. It is clear that $\mathfrak{\bar{a}}\subset \mathfrak{a}_{\bot}$. So the question is whether $\mathfrak{a}_{\bot}$ has additional roots, which are not the roots of $\mathfrak{\bar{a}}$ but are the linear combinations of them. It is possible when the set of angles between the roots of $\mathfrak{\bar{a}}$ does not contain all the angles between the roots of $\mathfrak{a}$, then reflecting the roots of $\mathfrak{a}$ by $s_{\alpha'}$ we get the additional roots of $\mathfrak{a}_{\bot}$.

All the cases are listed in the table \ref{tab:diagrams}.
\begin{table}[ph]
\label{tab:diagrams}
\noindent  \centering{
\begin{tabular}{|l|l|l|l|}
  \hline
  $\mathfrak{g}$ & Extended diagram of the algebra $\mathfrak{g}$ & Diagrams of the subalgebras $\mathfrak{a},\; \mathfrak{a}_{\bot}$ \\
  \hline
  $A_n$ & \includegraphics{A_series_ext_diag.pdf} & \includegraphics{A_series_ext_split.pdf} \\
  \hline
  $B_n$ & \includegraphics{B_series_ext_diag.pdf} & \includegraphics{B_series_ext_split.pdf} \\
  \hline
  $C_n$ & \includegraphics{C_series_ext_diag.pdf} & \includegraphics{C_series_ext_split.pdf} \\
  \hline
  $D_n$ & \includegraphics{D_series_ext_diag.pdf} & \includegraphics{D_series_ext_split.pdf} \\
  \hline
\end{tabular}
\caption{Subalgebras $\mathfrak{a},\; \mathfrak{a}_{\bot}$ for the classical series}}
\end{table}

For the algebra $\mathfrak{g}$ from the series $A_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r+1\}$ are $\Delta=\{\alpha_{ i j}= e_i- e_j,\; 1\leq i, j\leq r+1\}$, $\Delta^{+}=\{\alpha_{ij},\; i<j\}$ and the set of simple roots consists of $\alpha_{1,2},\alpha_{2,3},\dots,\alpha_{r,r+1}$. So for the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consisting of first $r_{\mathfrak{a}}$ simple roots we get $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{ i j},\; r_{\mathfrak{a}}+1<i,j\leq r+1\}$ and $\mathfrak{a}_{\bot}=A_{r-r_{\mathfrak{a}}-1}$.

For the algebra $\mathfrak{g}$ from the series $B_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r\}$ are $\Delta=\{\alpha_{\pm i,\pm j}=\pm e_i\pm e_j,\;i<j;\; \alpha_{\pm j}=\pm e_{ j},\; 1\leq j\leq r\}$, $\Delta^{+}=\{\alpha_{i,-j},\; \alpha_{ij},\; \alpha_j;\; i<j,\; 1\leq j\leq r\}$ and the set of simple roots consists of $\alpha_{1,-2},\alpha_{2,-3},\dots,\alpha_{r-1,-r},\alpha_r$. So if the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consists of first $r_{\mathfrak{a}}$ simple roots, then $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; r_{\mathfrak{a}}+1<i<j\leq r\}$ and $\mathfrak{a}_{\bot}=B_{r-r_{\mathfrak{a}}-1}$. Otherwise if $\mathfrak{a}=B_{r_{\mathfrak{a}}}$ and its simple roots are $\alpha_{r-r_{\mathfrak{a}}+1,-r+r_{\mathfrak{a}}-2},\dots,\alpha_{r-1,r},\alpha_r$ we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=B_{r-r_{\mathfrak{a}}}$. It is the only case when simple roots of $\mathfrak{a}_{\bot}$ can not be obtained from the extended Dynkin diagram, as can be seen in the Table \ref{tab:diagrams}. 
There exists the third possibility to get the pair of subalgebras $\mathfrak{a},\mathfrak{a}_{\bot}$ with the regular subalgebra $\mathfrak{a}$ by dropping single node from the extended Dynkin diagram of $B_r$. It can be done by taking as the set of simple roots of $\mathfrak{a}$ the set $\{\alpha_{1,-2},\alpha_{1,2},\alpha_{2,-3},\dots,\alpha_{r_{\mathfrak{a}}-1,-r_{\mathfrak{a}}}\}$. Then $\mathfrak{a}=D_{r_{\mathfrak{a}}}$, $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; r_{\mathfrak{a}}<i<j\leq r\}$ and $\mathfrak{a}_{\bot}=B_{r-r_{\mathfrak{a}}}$.

For the algebra $\mathfrak{g}$ from the series $C_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r\}$ are $\Delta=\{\alpha_{\pm i,\pm j}=\pm e_i\pm e_j,\;i<j;\; \alpha_{\pm j}=\pm 2e_{ j},\; 1\leq j\leq r\}$, $\Delta^{+}=\{\alpha_{i,-j},\; \alpha_{ij},\; \alpha_j;\; i<j,\; 1\leq j\leq r\}$ and the set of simple roots consists of $\alpha_{1,-2},\alpha_{2,-3},\dots,\alpha_{r-1,-r},\alpha_r$. So if the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consists of first $r_{\mathfrak{a}}$ simple roots, then $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; r_{\mathfrak{a}}+1<i<j\leq r\}$ and $\mathfrak{a}_{\bot}=C_{r-r_{\mathfrak{a}}-1}$. Otherwise if $\mathfrak{a}=C_{r_{\mathfrak{a}}}$ and its simple roots are $\alpha_{r-r_{\mathfrak{a}}+1,-r+r_{\mathfrak{a}}-2},\dots,\alpha_{r-1,r},\alpha_r$ we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=C_{r-r_{\mathfrak{a}}}$.

For the algebra $\mathfrak{g}$ from the series $D_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r\}$ are $\Delta=\{\alpha_{\pm i,\pm j}=\pm e_i\pm e_j,\;1\leq i<j\leq r\}$, $\Delta^{+}=\{\alpha_{i,-j},\; \alpha_{ij},\; i<j,\; 1\leq j\leq r\}$ and the set of simple roots consists of $\alpha_{1,-2},\alpha_{2,-3},\dots,\alpha_{r-1,-r},\alpha_{r-1,r}$. So if the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consists of first $r_{\mathfrak{a}}$ simple roots, then $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\; r_{\mathfrak{a}}+1<i<j\leq r\}$ and $\mathfrak{a}_{\bot}=D_{r-r_{\mathfrak{a}}-1}$. Otherwise if $\mathfrak{a}=D_{r_{\mathfrak{a}}}$ and its simple roots are $\alpha_{r-r_{\mathfrak{a}}+1,-r+r_{\mathfrak{a}}-2},\dots,\alpha_{r-1,r},\alpha_{r-1,r}$ we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=D_{r-r_{\mathfrak{a}}}$.

In the case of special embeddings the set $\Delta^{+}_{\bot}$ can be empty as for the
special embedding of $A_1\subset A_2$ with the embedding index equal to 4, or non-empty for example for the embedding $A_1\subset A_2\subset A_3$ which is depicted at the Figure \ref{fig:a1a2a3}. 
\begin{figure}[ph]
\noindent  \centering{
  \includegraphics[width=120mm]{special_A1_A2_A3.pdf}
  \caption{Special embedding $A_1\subset A_2\subset A_3$. Grey vectors are the roots of $A_2$, thick black - of $\mathfrak{a}=A_1$, dashed black are the orthogonal roots of $A_1$ which is contained in $\mathfrak{a}_{\bot}$ }
}
  \label{fig:a1a2a3}
\end{figure}

Using the existing classification of maximal special subalgebras \cite{dynkin1952semisimple} we immediately have the following pairs of the orthogonal subalgebras $\mathfrak{a},\;\mathfrak{a}_{\bot}$
\begin{equation}
  \label{eq:42}
  \begin{array}{lll}
      su(p)\oplus su(q) & \subset su(pq) &\\
      so(p)\oplus so(q) & \subset so(pq) &\\
      sp(2p)\oplus sp(2q) & \subset so(4pq)&\\
      sp(2p)\oplus so(q) & \subset sp(2pq)&\\
      so(p)\oplus so(q) & \subset so(p+q)& \text{for}\;p\;\text{and}\;q\;\text{odd}\\
  \end{array}
\end{equation}
Exceptional Lie algebras and other non-maximal subalgebras will be considered elsewhere.

Up to this point we considered the problem of $\mathfrak{a}_{\bot}$-construction given $\mathfrak{a}\in\mathfrak{g}$ for regular injections in terms of Dynkin diagrams. When the root systems $ \Delta$ and $\Delta_{\mathfrak{a}}$ are known explicitly all that we need is to select the roots  
$\Delta_{\bot}=\left\{\alpha\in \Delta:\alpha\bot \Delta_{\mathfrak{a}}\right\}$ and correspondingly the positive roots
$\Delta^{+}_{\bot}=\left\{\alpha\in \Delta^{+}:\alpha\bot \Delta_{\mathfrak{a}}\right\}$.

Now consider the $\mathfrak{a}_{\bot}\oplus\mathfrak{h}_{\mathfrak{a}}$-module 
with the highest weight $\mu$. For the character of this module we have
\begin{equation}
  \label{eq:41}
  ch L^{\mu}_{\mathfrak{a}_{\bot}\oplus \mathfrak{h}}= \frac{\sum_{\omega\in W_{\bot}} \epsilon(\omega) e^{\omega(\mu+\rho_{\mathfrak{a}_{\bot}})-\rho_{\mathfrak{a}_{\bot}}}}{\prod_{\alpha\in\Delta^{+}_{\bot}}(1-e^{-\alpha})^{\mathrm{mult}(\alpha)}}
\end{equation}
The projection $\pi_{\mathfrak{a}}(ch\, L^{\mu}_{\mathfrak{a}_{\bot}\oplus \mathfrak{h}})$ gives us the single element $e^{\pi_{\mathfrak{a}} \cdot\mu}$ of the formal algebra ${\cal E}\left( \mathfrak{a}\right)$ with the multiplicity equal to the dimension of the module $L^{\mu}_{\mathfrak{a}_{\bot}\oplus \mathfrak{h}}$, since all the roots of $\mathfrak{a}_{\bot}$ are orthogonal to  that of $\Delta_{\mathfrak{a}}$.

>>>>

\begin{document}
\title{\textbf{{\Large {System of dual functions in the reduction of affine Lie algebra representations}}}}
\author{Vladimir Lyakhovsky \thanks{ Supported by
 RFFI grant N 09-01-00504 and the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:lyakh1507@nm.ru \\
[5mm] Anton Nazarov \thanks{ Supported by
the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:antonnaz@gmail.com
}
\maketitle

\begin{abstract}
  Using the recurrent relation for the series expansion of the affine Lie algebra branching functions we construct the dual set of the functions, which can be completely described using Weyl symmetry of affine Lie algebra. Explicit construction of the dual functions can be carried out using the geometric finite automata approach. The task of the explicit construction of the branching functions is then rewritten as the procedure of the inversion of matrix of dual functions. Finally we study the properties of the dual functions under modular and conformal transformations and discuss possible applications in two-dimensional conformal field theory.
\end{abstract}
